---
title: "Course Project 1 - Reproducible Research"
author: "Henry lopez"
date: "25 de julio de 2020"
output: html_document
---

#Assignment Instructions

1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


##Step 1
##Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
rm(list=ls())
# Download archive file, if it does not exist

archiveFile <- "repdata_data_activity"

if(!file.exists(archiveFile)) {
  
  archiveURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  
  download.file(url=archiveURL,destfile=archiveFile)
}

if(!(file.exists("activity.csv"))){ 
  unzip(archiveFile) 
}

#Loading and preprocessing the data
data <- read.csv("activity.csv")
str(data)
data$date <- as.Date(data$date)
data1 <- na.omit(data)
```



##Step 2
##Histogram of the total number of steps taken each day
```{r, echo = TRUE}
library(dplyr)
steps_per_day <- data1 %>% 
                 group_by(date) %>% 
                  summarise(sumsteps =sum(steps,na.rm=T))
#Make a histogram of the total number of steps taken each day.

png("plot1.png")
hist(steps_per_day$sumsteps,
     main="Histogram of steps per day",
     xlab = "Steps",ylab="Frecuency",
     breaks = seq(0,25000,2500),ylim = c(0,20),
     col="blue")
dev.off()
hist(steps_per_day$sumsteps,
     main="Histogram of steps per day",
     xlab = "Steps",ylab="Frecuency",
     breaks = seq(0,25000,2500),ylim = c(0,20),
     col="blue")


```

##Step 3
##Mean and median number of steps taken each day
```{r, echo = TRUE}
mean_steps <- mean(steps_per_day$sumsteps);mean_steps
median_steps <- median(steps_per_day$sumsteps);median_steps
```


##Step 4
##Time series plot of the average number of steps taken
```{r, echo = TRUE}
stepsPerInterval <- data1 %>% 
                    group_by(interval) %>% 
                    summarise(meansteps=mean(steps,na.rm=T))

png("plot2.png")

plot(stepsPerInterval$meansteps~stepsPerInterval$interval,
     col="black",type="l", main="Steps By Time Interval",
     xlab = "Intervals", ylab = "Average Number of Steps")
dev.off()

plot(stepsPerInterval$meansteps~stepsPerInterval$interval,
     col="black",type="l", main="Steps By Time Interval",
     xlab = "Intervals", ylab = "Average Number of Steps")

```


##Step 5
##The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo = TRUE}

print(paste("Interval containing the most steps on average: ",stepsPerInterval$interval[stepsPerInterval$meansteps==max(stepsPerInterval$meansteps)]))

print(paste("Interval containing the most steps on average: ",round(stepsPerInterval$meansteps[stepsPerInterval$meansteps==max(stepsPerInterval$meansteps)])),digits = 4)



```


##Step 6
```{r, echo = TRUE}

sum(is.na(data$steps))


#So, let's create a new dataset that is equal to the original dataset but with the missing data filled in.

index<-which(is.na(data$steps))
l<-length(index)
steps_avg<-with(data,tapply(steps,date,mean,na.rm=TRUE))
na<-mean(steps_avg,na.rm = TRUE)
for (i in 1:l) {
  data[index[i],1]<-na
}



#Let's see if we filled all NAs properly and see how our new dataset looks like
sum(is.na(data$steps))


str(data)

#hist

steps_2<-with(data,tapply(steps,date,sum,na.rm=TRUE))

png("plot3.png")
hist(steps_2,col="blue",breaks = seq(0,25000,2500),ylim = c(0,20), main = "Total number of steps taken each day (filled data)",xlab = "Number of steps per day")
dev.off()
hist(steps_2,col="blue",breaks = seq(0,25000,2500),ylim = c(0,20), main = "Total number of steps taken each day (filled data)",xlab = "Number of steps per day")
print(mean_steps_2<-mean(steps_2))
print(median_steps_2<-median(steps_2))
```  

## Step 7
Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo = TRUE}
# Create variable with date in correct format
data1$RealDate <- as.Date(data1$date, format = "%Y-%m-%d")
# create a variable with weekdays name
data1$weekday <- weekdays(data1$RealDate)
# create a new variable indicating weekday or weekend
data1$DayType <- ifelse(data1$weekday=='sabado' | data1$weekday=='domingo', 'weekend','weekday')
# see first 10 values
head(data1, n=10)


# create table with steps per time across weekdaydays or weekend days
StepsPerTimeDT <- aggregate(steps~interval+DayType,data=data1,FUN=mean,na.action=na.omit)
StepsPerTimeDT$time <- stepsPerInterval$interval/100

library(ggplot2)

png("plot4.png")
j <- ggplot(StepsPerTimeDT, aes(time, steps))
j+geom_line(col="darkred")+ggtitle("Average steps per time interval: weekdays vs. weekends")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))+facet_grid(DayType ~ .)
dev.off()


j <- ggplot(StepsPerTimeDT, aes(time, steps))
j+geom_line(col="darkred")+ggtitle("Average steps per time interval: weekdays vs. weekends")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))+facet_grid(DayType ~ .)
```  




